---
title: "Statistical Inference Project Part 2"
author: "Oliver Hofkens"
date: "5/26/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1: Basic Inferential Data Analysis

In this report, we will compare the length of odontoblasts (cells responsible for tooth growth)
in 60 guinea pig by supplement and dose. The data comes from the ToothGrowth dataset in R.

### Data Summary & Exploratory Analysis 

We load the data and take a quick look at it:

```{r}
data <- ToothGrowth
str(data)
```
```{r}
summary(data)
```


```{r message=FALSE, warning=FALSE}
library(ggplot2)

g <- ggplot(data, aes(dose, len, colour = supp)) +
    geom_point() +
    geom_smooth()
g
```

This scatterplot (with overlaid smoothed means) suggests that a higher dose will 
result in bigger length, and that orange juice has a bigger impact than vitamin C 
at lower doses.

### Hypothesis Testing

#### 1. Does OJ stimulate bigger length than VC at doses lower than 1mg?

```{r}
lowdoses <- subset(data, dose <= 1)
t.test(len ~ supp ,lowdoses, alternative = "greater")
```
A p-value of 0.0021 means we can confidently reject the null hypothesis that 
there is no real difference in length between 2 supplements at low doses, and we 
accept our alternative hypothesis that OJ stimulates bigger length at doses lower 
than 1mg. 

#### 2. Does a higher dose of any supplement stimulate bigger length?

##### 2a. Doses of 1mg and smaller
```{r}
t.test(len ~ dose, lowdoses, alternative = "less")
```

Again, such a small P-value means we can confidently say that we're right, bigger 
doses stimulate bigger length.

#### 2b. Doses of 1mg and up

```{r}
highdoses <- subset(data, dose >= 1)
t.test(len ~ dose, highdoses, alternative = "less")
```

Again, such a small P-value means we can confidently say that we're right, bigger 
doses stimulate bigger length.

### Conclusions

* A higher dose (independent of supplement) results in bigger length.
* In doses of 0.5 and 1 mg, OJ shows a bigger increase in length than VC.

### Assumptions

* The guinea pigs in this experiment were randomly picked from a population.
* Each observation in the dataset is a different guinea pig.
