---
title: "Statistical Inference Project Part 1"
author: "Oliver Hofkens"
date: "25 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1: Simulation Exercise

In this report, 1000 random exponentials are simulated and the distribution
of their averages is investigated. The sample mean and variance are compared to 
the theoretical statistics, and it is shown that the distribution is approximately
normal.

### Simulations

Let's start by taking a look at the general form of an exponential.
```{r exp}
library(ggplot2)

set.seed(123456789)
lambda <- 0.2

exampleExp <- rexp(1000, rate = lambda)
exampleExpStats <- summary(exampleExp)
exampleExp <- data.frame(exampleExp)

g <- ggplot(exampleExp, aes(exampleExp)) + 
    geom_histogram() + 
    geom_vline(aes(xintercept = mean(exampleExp), color = 'mean')) + 
    geom_vline(aes(xintercept = sd(exampleExp), color = 'stddev')) +
    xlab('value') + 
    scale_x_continuous(breaks = seq(0, 30, 5))
g
```

We can see that the mean and standard deviation are around 5 for a lambda of 0.2, 
which fits the theoretical values (1/lambda).

```{r params}
nSimulations <- 1000
```

```{r theoretical}
mu0 <- 1/lambda
sd0 <- 1/lambda
```
