---
title: "Statistical Inference Project Part 1"
author: "Oliver Hofkens"
date: "25 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1: Simulation Exercise

In this report, 1000 random exponentials are simulated and the distribution
of their averages is investigated. The sample mean and variance are compared to 
the theoretical statistics, and it is shown that the distribution is approximately
normal.

### Simulations

Let's start by taking a look at the general form of an exponential.
```{r exp, message=FALSE}
library(ggplot2)

set.seed(123456789)
lambda <- 0.2

exampleExp <- rexp(1000, rate = lambda)
exampleExpStats <- summary(exampleExp)
exampleExp <- data.frame(exampleExp)

g <- ggplot(exampleExp, aes(exampleExp)) + 
    geom_histogram() + 
    geom_vline(aes(xintercept = mean(exampleExp), color = 'mean')) + 
    geom_vline(aes(xintercept = sd(exampleExp), color = 'stddev')) +
    xlab('value') + 
    scale_x_continuous(breaks = seq(0, 30, 5))
g
```

We can see that the mean and standard deviation are `r mean(exampleExp$exampleExp)` and `r sd(exampleExp$exampleExp)` for a lambda of `r lambda`, 
which fits the theoretical values (1/lambda).

Let's generate 1000 random exponentials of 40 samples each for our simulation:
```{r params}
nSimulations <- 1000
nSamples <- 40

# Simulate 1000 random exponentials of 40 samples
sims <- replicate(nSimulations, rexp(nSamples, rate = lambda))
```

Then let's calculate the mean, standard deviation, and variance and plot the 
distribution of sample means of the simulations.

```{r simdist, message=FALSE}
# calculate the mean of each simulation
means <- apply(sims, 2, mean)
mua <- mean(means)
sda <- sd(means) 

means <- data.frame(means)

g <- ggplot(means, aes(means)) + 
    geom_histogram(aes(y = ..density..)) +
    geom_vline(aes(xintercept = mua, color = 'mean')) + 
    geom_vline(aes(xintercept = mua + sda, color = 'stddev')) +
    geom_vline(aes(xintercept = mua - sda, color = 'stddev')) +
    stat_function(fun = dnorm, args = list(mean = mua, sd = sda)) +
    scale_x_continuous(breaks = 1:10)
g
```

### Mean: Sample vs Theoretical

On first sight we can see that the distribution of sample means looks normal, with the mean at `r mua`. 

```{r theoreticalmean}
mu0 <- 1/lambda
diff <- abs(mu0 - mua)
```

The theoretical mean is `r mu0`, so the mean of sample means is only off by `r diff`.

### Variance: Sample vs Theoretical

```{r theoreticalvariance} 
sd0 <- 1/lambda/sqrt(nSamples)
var0 <- sd0^2

vara <- var(means)

diff <- abs(var0 - vara)
```

The theoretical variance is `r var0` and our simulation variance is `r vara`, so the variance of sample means is only off by `r diff`.
